### 2024-01-31 supplementary_materials/supplementary_materials.Rmd
start supplementary materials

---
title: "Supplementary Materials"
author:
  - name: Michael Schramm
    affiliation: '1'
    etal: Schramm # First author's last name. 
    email: michael.schramm@ag.tamu.edu # Indicates corresponding Author
    institution: Texas Water Resources Institute, Texas A&M AgriLife Research
    street: 1001 Holleman Dr E.
    city: College Station
    state: Texas # only USA, Australia, Canada
    zip: 77840
    country: USA
  - name: Duncan Kikoyo
    affiliation: '1'
  - name: Janelle Wright
    affiliation: '1'
  - name: Shubham Jain
    affiliation: '1'
affiliation:
  - id: '1'
    department: Texas Water Resources Institute
    institution: Texas A&M AgriLife Research
    city: College Station
    state: Texas # only USA, Australia, Canada
    country: USA
output: 
  bookdown::pdf_book:
    template: template.tex
# bib yaml is commented to manually insert in correct section
# csl: frontiers.csl
# bibliography: bibliography.bib
# Match bibliostyle with the documentclass
# biblio-style: Frontiers-Harvard # Many Frontiers journals use the Harvard referencing system (Author-date), to find the style and resources for the journal you are submitting to: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal. For Humanities and Social Sciences articles please include page numbers in the in-text citations 
# biblio-style: Frontiers-Vancouver # Many Frontiers journals use the numbered referencing system, to find the style and resources for the journal you are submitting to: https://zendesk.frontiersin.org/hc/en-us/articles/360017860337-Frontiers-Reference-Styles-by-Journal
---

```{r setup, include=FALSE}
## load libraries
library(ggtext)
library(kableExtra)
library(officer)
library(ragg)
library(targets)
library(tidyverse)

## this sets our default code chunk options
knitr::opts_chunk$set(dev = "ragg_png",
                      echo = FALSE,
                      ## figure chunk options
                      fig.cap = TRUE,
                      fig.width = 6,
                      dpi = 300)

## make fonts available
font_hoist <- function(family, silent = FALSE) {
  font_specs <- systemfonts::system_fonts() |> 
    filter(family == {{family}}) |> 
    mutate(family = paste(family, style)) |> 
    select(plain = path, name = family)

  purrr::pwalk(as.list(font_specs), systemfonts::register_font)

  if (!silent)  message(paste0("Hoisted ", nrow(font_specs), " variants:\n",
                               paste(font_specs$name, collapse = "\n")))
}

font_hoist("Manrope")

## grab output from targets to use in manuscript
store <- "../_targets"

```

```{r bacsummary}

tar_read(bac_table, store = store) |> 
  pack_rows(group_label = "BMP Subcategories",
            start_row = 3,
            end_row = 5) |> 
  pack_rows(group_label = "Arididty:BMP Subcategory Interaction",
            start_row = 7,
            end_row = 9)

```

```{r tnsummary}

tar_read(tn_table, store = store)

```

```{r dinsummary}

tar_read(tin_table, store = store)

```

```{r tpsummary}

tar_read(tp_table, store = store)

```

```{r po4summary}

tar_read(po4_table, store = store)

```

```{r tsssummary}

tar_read(tss_table, store = store)

```

```{r bacpub}
#| fig.cap: "Plot of adjusted sampling error and predicted effect size for FIB multilevel random effects model with moderators. Slope and 95% confidence interval of the adjusted sampling error term is annoted."
#| fig.height: 3
#| fig.width: 6
tar_read(bac_sse, store = store)$plot
```

```{r tnpub}
#| fig.cap: "Plot of adjusted sampling error and predicted effect size for TN multilevel random effects model with moderators. Slope and 95% confidence interval of the adjusted sampling error term is annoted."
#| fig.height: 3
#| fig.width: 6
tar_read(tn_sse, store = store)$plot
```

```{r dinpub}
#| fig.cap: "Plot of adjusted sampling error and predicted effect size for DIN multilevel random effects model with moderators. Slope and 95% confidence interval of the adjusted sampling error term is annoted."
#| fig.height: 3
#| fig.width: 6
tar_read(tin_sse, store = store)$plot
```

```{r tppub}
#| fig.cap: "Plot of adjusted sampling error and predicted effect size for TP multilevel random effects model with moderators. Slope and 95% confidence interval of the adjusted sampling error term is annoted."
#| fig.height: 3
#| fig.width: 6
tar_read(tp_sse, store = store)$plot
```

```{r po4pub}
#| fig.cap: "Plot of adjusted sampling error and predicted effect size for PO\\textsubscript{4} multilevel random effects model with moderators. Slope and 95% confidence interval of the adjusted sampling error term is annoted."
#| fig.height: 3
#| fig.width: 6
tar_read(po4_sse, store = store)$plot
```

```{r tsspub}
#| fig.cap: "Plot of adjusted sampling error and predicted effect size for TSS multilevel random effects model with moderators. Slope and 95% confidence interval of the adjusted sampling error term is annoted."
#| fig.height: 3
#| fig.width: 6
tar_read(tn_sse, store = store)$plot
```

```{r bacsens}
#| fig.cap: "Plot of intercept estimates from sensitivity analysis for the FIB regression model. Individual points are the intercept estimates with 95% confince intervals for the regression model fit leaving out the study value indicated on the y-axis. The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values. Any study values outside of the full model confidence intervals would be considered and outlier."
#| fig.height: 8
#| fig.width: 8
#| out.width: "100%"
tar_read(bac_sens, store = store)$p1
```

```{r tnsens}
#| fig.cap: "Plot of intercept estimates from sensitivity analysis for the TN regression model. Individual points are the intercept estimates with 95% confince intervals for the regression model fit leaving out the study value indicated on the y-axis. The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values. Any study values outside of the full model confidence intervals would be considered and outlier."
#| fig.height: 5.5
#| fig.width: 8
#| out.width: "100%"
tar_read(tn_sens, store = store)$p1
```

```{r dinsens}
#| fig.cap: "Plot of intercept estimates from sensitivity analysis for the DIN regression model. Individual points are the intercept estimates with 95% confince intervals for the regression model fit leaving out the study value indicated on the y-axis. The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values. Any study values outside of the full model confidence intervals would be considered and outlier."
#| fig.height: 7
#| fig.width: 8
#| out.width: "100%"
tar_read(tin_sens, store = store)$p1
```

```{r tpsens}
#| fig.cap: "Plot of intercept estimates from sensitivity analysis for the TP regression model. Individual points are the intercept estimates with 95% confince intervals for the regression model fit leaving out the study value indicated on the y-axis. The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values. Any study values outside of the full model confidence intervals would be considered and outlier."
#| fig.height: 5.5
#| fig.width: 8
#| out.width: "100%"
tar_read(tp_sens, store = store)$p1
```

```{r po4sens}
#| fig.cap: "Plot of intercept estimates from sensitivity analysis for the PO\\textsubscript{4} regression model. Individual points are the intercept estimates with 95% confince intervals for the regression model fit leaving out the study value indicated on the y-axis. The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values. Any study values outside of the full model confidence intervals would be considered and outlier."
#| fig.height: 4
#| fig.width: 8
#| out.width: "100%"
tar_read(po4_sens, store = store)$p1
```

```{r tsssens}
#| fig.cap: "Plot of intercept estimates from sensitivity analysis for the TSS regression model. Individual points are the intercept estimates with 95% confince intervals for the regression model fit leaving out the study value indicated on the y-axis. The vertical solid line and dotted lines indicate the intercept and 95% confidence intervals for the model that includes all values. Any study values outside of the full model confidence intervals would be considered and outlier."
#| fig.height: 4.5
#| fig.width: 8
#| out.width: "100%"
tar_read(tss_sens, store = store)$p1
```

```{r bacforest}
#| fig.cap: "Forest plot of FIB effect size estimates and 95% confidence intervals. Size of points are scaled to inverse of the sampling variance."
#| fig.height: 8
#| fig.width: 8
#| out.width: "100%"
tar_read(bac_forest_plot, store = store)
```

```{r tnforest}
#| fig.cap: "Forest plot of TN effect size estimates and 95% confidence intervals. Size of points are scaled to inverse of the sampling variance."
#| fig.height: 8
#| fig.width: 8
#| out.width: "100%"
tar_read(tn_forest_plot, store = store)
```

```{r dinforest}
#| fig.cap: "Forest plot of DIN effect size estimates and 95% confidence intervals. Size of points are scaled to inverse of the sampling variance."
#| fig.height: 8
#| fig.width: 8
#| out.width: "100%"
tar_read(tin_forest_plot, store = store)
```

```{r tpforest}
#| fig.cap: "Forest plot of TP effect size estimates and 95% confidence intervals. Size of points are scaled to inverse of the sampling variance."
#| fig.height: 8
#| fig.width: 8
#| out.width: "100%"
tar_read(tp_forest_plot, store = store)
```

```{r po4forest}
#| fig.cap: "Forest plot of PO\\textsubscript{4} effect size estimates and 95% confidence intervals. Size of points are scaled to inverse of the sampling variance."
#| fig.height: 8
#| fig.width: 8
#| out.width: "100%"
tar_read(bac_forest_plot, store = store)
```

```{r tssforest}
#| fig.cap: "Forest plot of TSS effect size estimates and 95% confidence intervals. Size of points are scaled to inverse of the sampling variance."
#| fig.height: 8
#| fig.width: 8
#| out.width: "100%"
tar_read(tss_forest_plot, store = store)
```